Hamby Data and `rtrees`

### 2019-11-08 journals/01-hamby_data_and_models/01-hamby_data_and_models.Rmd
created some partial dependence plots

---
title: "Hamby Data and `rtrees`"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: textmate
---

@@

```{r}
# Load libraries
library(bulletr)
library(cowplot)
library(GGally)
library(glmnet)
library(plotly)
library(randomForest)
library(tidyverse)
```

@@

```{r}
# Create the plot to look at number of comparisons within the known bullets
countplot <- hamby173and252_raw %>%
  filter(barrel1 %in% c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
         barrel2 %in% c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")) %>%
  group_by(study1, study2, barrel1, barrel2, bullet1, bullet2) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = barrel1, y = barrel2)) +
  geom_tile(aes(fill = count)) + 
  facet_grid(study1 + bullet1 ~ study2 + bullet2, scales = "free") +
  theme_minimal() + 
  scale_fill_distiller(palette = "GnBu", direction = 1)

@@

```{r fig.width = 10}
hamby173and252_train %>%
  mutate(case = 1:n()) %>%
  select(case, rfscore, samesource, ccf:sum_peaks) %>%
  mutate_at(rf_features, scale) %>%
  gather(variable, value, 4:12) %>%
  mutate(variable = factor(variable, levels = rf_features_ordered$feature)) %>%
  ggplot(aes(x = variable, y = value, group = case, color = rfscore)) + 
  geom_point() + 
  geom_line(alpha = 0.2) + 
  facet_grid( ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values")
```

@@

```{r fig.width = 10}
hamby224_test %>%
  na.omit() %>%
  select(case, set, rfscore, samesource, ccf:sum_peaks) %>%
  mutate_at(rf_features, scale) %>%
  gather(variable, value, 5:13) %>%
  mutate(variable = factor(variable, levels = rf_features_ordered$feature)) %>%
  ggplot(aes(x = variable, y = value, group = case, color = rfscore)) + 
  geom_point() + 
  geom_line(alpha = 0.5) + 
  facet_grid(set ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values")
```

>>>>

---
title: "Hamby Data and `rtrees`"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

@@

```{r}
# Load libraries
library(bulletr)
library(cowplot)
library(ggConvexHull) #devtools::install_github("cmartin/ggConvexHull")
library(GGally)
library(ggpcp)
library(glmnet)
library(plotly)
library(randomForest)
library(tidyverse)
```

@@

```{r}
# Create the plot to look at number of comparisons within the known bullets
countplot <- hamby173and252_raw %>%
  filter(barrel1 %in% c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10"),
         barrel2 %in% c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10")) %>%
  group_by(study1, study2, barrel1, barrel2, bullet1, bullet2) %>%
  summarise(count = n()) %>%
  ggplot(aes(x = barrel1, y = barrel2)) +
  geom_tile(aes(fill = count)) + 
  facet_grid(study1 + bullet1 ~ study2 + bullet2, scales = "free") +
  theme_minimal() + 
  scale_fill_distiller(palette = "GnBu", direction = 1) + 
  theme(legend.position = "bottom")

@@

```{r fig.width = 14}
hamby173and252_train %>%
  na.omit() %>%
  ggplot(aes(color = rfscore)) + 
  geom_pcp(aes(vars = vars(as.vector(rf_features_ordered$feature))), alpha = 0.1) + 
  facet_grid(. ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values") 
```

@@

```{r fig.width = 14}
hamby224_test %>%
  na.omit() %>%
  ggplot(aes(color = rfscore)) + 
  geom_pcp(aes(vars = vars(as.vector(rf_features_ordered$feature))), alpha = 0.1) + 
  facet_grid(. ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values") 
```

##### Partial Dependence Plots

I tried using the package `pdp`, but I got the following error.

```{r error = TRUE}
library(pdp)
partial(object = rtrees, pred.var = "rough_cor", plot = TRUE, plot.engine = "ggplot2")
```

Then I tried creating some plots myself.

```{r}
# Specify the number of breaks in the grid
n <- 100000

# Create a prediction grid
set.seed(20191108)
simulated_data <- hamby173and252_train %>%
  select(rf_features) %>%
  gather(variable, value) %>%
  group_by(variable) %>%
  summarise(min = min(value),
            max = max(value)) %>%
  rowwise() %>%
  mutate(samples = list(runif(n, min, max))) %>%
  select(variable, samples) %>%
  unnest() %>%
  mutate(case = rep(1:n, 9)) %>%
  spread(variable, samples) %>%
  select(-case)

simulated_data$rfpred <- predict(rtrees, newdata = simulated_data, type = "prob")[,2]
```

```{r}
simulated_data %>%
  gather(key = variable, value = value, rf_features) %>%
  ggplot(aes(x = value, y = rfpred, color = rfpred)) + 
  geom_point() + 
  facet_wrap(. ~ variable, scales = "free_x") + 
  scale_color_viridis_c("rfpred", option = "A", limits = c(0, 1))
```

```{r}
ggplot() + 
  stat_summary_2d(data = simulated_data,
                  mapping = aes(x = ccf, y = rough_cor, z = rfpred), 
                  bins = 75, 
                  fun = mean) + 
  scale_fill_viridis_c("rfpred", option = "A", limits = c(0, 1)) + 
  geom_convexhull(data = hamby173and252_train, 
                  mapping = aes(x = ccf, y = rough_cor), 
                  alpha = 0, 
                  color = "black")
```

```{r}
ggplot() + 
  stat_summary_2d(data = simulated_data,
                  mapping = aes(x = ccf, y = matches, z = rfpred), 
                  bins = 75, 
                  fun = mean) + 
  scale_fill_viridis_c("rfpred", option = "A", limits = c(0, 1)) + 
  geom_convexhull(data = hamby173and252_train, 
                  mapping = aes(x = ccf, y = matches), 
                  alpha = 0, 
                  color = "black")
```
